Towards Compiler-Agnostic Performance in Finite-Difference Codes

### 2015-09-14 docs/papers/shallow_performance/shallow_perf.tex
parco paper: finish new para describing Gnu slowdown in going from 256 to 512.

The primary aim of this work is to determine whether any performance
loss due to the \psykal\ separation of concerns can be negated by
optimising the PSy layer and if so, what the key transformations are
(for a given system). To evaluate this we perform tests on a range of
hardware and a range of compilers. The hardware and compilers are
listed in Table~\ref{TABLE_compilers}. Where a compiler is available
on a given piece of hardware, the version number used in this work is
specified.
%
The Intel Haswell-based system (Xeon E5-1620 v2) has a clock speed of
3.7~GHz and 10~MB of last-level cache. The Intel Ivy Bridge-based
system (Xeon E5-2697) has a clock speed of 2.7~GHz and a last-level
cache of 30~MB. The CPU in the IBM Power-8 system is built around 12
`chiplets'. Each chiplet has a single core with 64~KB of L1 cache,
512~KB of L2 cache and 8~MB of L3 cache. The L3 cache of each chiplet
is shared with all other chiplets so that there is a total of 96~MB of
last-level cache. The cores on the system we used had a clock speed of
3.3~GHz.

@@

In Figure~\ref{FIG_orig_perf_summary} we plot the performance of the
original version of the Shallow code for the range of compilers and
hardware considered here. This summary demonstrates the effect of the
larger last-level cache of the Intel Ivy Bridge system compared to
that of the Haswell system; note the drop in performance
in going from a problem size of $256^{2}$ to a problem size of
$512^{2}$ for the first two bars. The performance of the Ivy
Bridge system with the Cray or Intel compiler (penultimate two bars)
only drops to this level when the domain size is increased to
$1024^{2}$. At this point, the working set no longer fits within cache
and performance is dominated by the bandwidth to main memory, making
it relatively insensitive to the choice of compiler.

@@

Strangely, the third bar (Gnu compiler) in
Figure~\ref{FIG_orig_perf_summary} shows the same drop in performance
in moving from the $256^2$ case to the $512^2$ case, despite the
larger last-level cache of the Ivy Bridge system. In order to
investigate this further, we profiled the application to obtain
hardware performance counter data. This revealed that the drop in
performance is due to a XXX

 With the
Intel compiler this drop in efficiency only occurs when going to
$512^2$. We therefore conclude that the drop in performance is due to
the Gnu compiler generating code that YYY
compared to that produced by the Intel and Cray
compilers.

@@

Finally, we compare the performance of the best \psykal\ version of
Shallow with that of the original. In
Figure~\ref{FIG_slowdown_summary} we plot the percentage {\em
  difference} between the performance of the original and the (best)
\psykal\ versions of Shallow for each compiler/CPU combination. The
most important feature of this plot is that the performance penalty
incurred by the \psykal\ version is less than four percent with only
one exception where it is 6.6\%.  In fact, for the Intel and Cray
compilers, the \psykal\ version of Shallow is never more than 2\%
slower than the original and in some cases is faster.  This
demonstrates that we can reliably recover the performance of the
original version of the code, despite the significant restructuring
required by the \psykal\ approach.

>>>>

The primary aim of this work is to determine whether any performance
loss due to the \psykal\ separation of concerns can be negated by
optimising the PSy layer and if so, what the key transformations are
(for a given system). To evaluate this we perform tests on a range of
hardware and a range of compilers. The hardware and compilers are
listed in Table~\ref{TABLE_compilers}. Where a compiler is available
on a given piece of hardware, the version number used in this work is
specified.
%
The Intel Haswell-based system (Xeon E5-1620 v2) has a clock speed of
3.7~GHz and 10~MB of last-level cache (LLC). The Intel Ivy
Bridge-based system (Xeon E5-2697) has a clock speed of 2.7~GHz and a
LLC of 30~MB. The CPU in the IBM Power-8 system is built around 12
`chiplets'. Each chiplet has a single core with 64~KB of L1 cache,
512~KB of L2 cache and 8~MB of L3 cache. The L3 cache of each chiplet
is shared with all other chiplets so that there is a total of 96~MB of
LLC. The cores on the system we used had a clock speed of 3.3~GHz.

@@

In Figure~\ref{FIG_orig_perf_summary} we plot the performance of the
original version of the Shallow code for the range of compilers and
hardware considered here. This summary demonstrates the effect of the
larger LLC of the Intel Ivy Bridge system compared to
that of the Haswell system; note the drop in performance
in going from a problem size of $256^{2}$ to a problem size of
$512^{2}$ for the first two bars. The performance of the Ivy
Bridge system with the Cray or Intel compiler (penultimate two bars)
only drops to this level when the domain size is increased to
$1024^{2}$. At this point, the working set no longer fits within cache
and performance is dominated by the bandwidth to main memory, making
it relatively insensitive to the choice of compiler.

@@

Strangely, the third bar (Gnu compiler) in
Figure~\ref{FIG_orig_perf_summary} shows the same drop in performance
in moving from the $256^2$ case to the $512^2$ case, despite the
larger LLC of the Ivy Bridge system. In order to investigate this
further, we profiled the application to obtain hardware performance
counter data for the Gnu- and Intel-compiled executables. This
revealed that the Gnu-compiled executable made more references to
memory and had a higher LLC miss-rate compared with the Intel binary.
We conclude that the Gnu compiler is not utilising registers as
efficiently as the Intel compiler and that this results in greater
memory traffic.

@@

Finally, we compare the performance of the best \psykal\ version of
Shallow with that of the original. In
Figure~\ref{FIG_slowdown_summary} we plot the percentage
  difference between the performance of the original and the (best)
\psykal\ versions of Shallow for each compiler/CPU combination. The
most important feature of this plot is that the performance penalty
incurred by the \psykal\ version is less than four percent with only
one exception where it is 6.6\%.  In fact, for the Intel and Cray
compilers, the \psykal\ version of Shallow is never more than 2\%
slower than the original and in some cases is faster.  This
demonstrates that we can reliably recover the performance of the
original version of the code, despite the significant restructuring
required by the \psykal\ approach.
